---
title: "DATA607 Assignment #3:\n R Character Manipulation"
author: "Bonnie Cooper"
date: "2/16/2020"
output:
  html_document:
    highlight: tango
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br><br>  

## 1. College Majors Dataset
<font size="4">
Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"


```{r , message=F}
#load the FiveThirtyEight data from 'majors-list.csv' to a data.frame
library( tidyverse )
dataURL <- 'https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv'
dataDF <- read_csv( dataURL )
head( dataDF )
```
```{r}
#identify majors that contain the word 'DATA'
DATA_idx <- grepl( 'DATA', dataDF$Major )
dataDF$Major[ DATA_idx ]
```
```{r}
#identify majors that contain the word 'STATISTICS'
DATA_idx <- grepl( 'STATISTICS', dataDF$Major )
dataDF$Major[ DATA_idx ]
```
```{r}
#I'm just curious. See a lot of majors with 'SCIENCE' or 'SCIENCES'. 
#Would like to identify & count them as well:
DATA_idx <- grepl( 'SCIENCE | SCIENCES', dataDF$Major )
dataDF$Major[ DATA_idx ]
length( dataDF$Major[ DATA_idx ] )
```
<br><br>

----------------------------------------------------------------------------



## 2. Parse words to list

<font size="4">
Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}
#make character vector that hold desired output to use as a test case:
testStrings <- c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

#I was really confused by the wording of this question 
#& am still not sure what we are supposed to work with as input. 
#This was my best guess:
data <- 
'[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"'
data
```
<br>  
We want to parse the words from the string above and pass them to a character vector using regular expressions
```{r}
#use strsplit() to split by '\"'
result <- strsplit(data, '\\"')
# '[a-z].*[a-z]' -> find any lower case alphabet, 
#there can be any number of characters in between, 
#it has the end with a lower case alphabet number.
dataStrings = grep(pattern = '[a-z].*[a-z]', result[[1]], value = TRUE, ignore.case = TRUE)
dataStrings
#test to see if our result matches our known test character vector
dataStrings == testStrings
```
<br><br>

----------------------------------------------------------------------------
  


## 3. Describe, in words, what these expressions will match:
<font size="4">


* ### (.)\1\1  ->    
'(.)' Find a character '\1' that repeats and '\1' then repeats again ->  
find 3 repreating characters (e.g. 'aaa')  
<br> 

* ### (.)(.)\\2\\1 ->     
'(.)'=Find a character '(.)'= Find the next character '\\2'= then the have the second character '\\1'= then the contents of the 1st character ->
Find a four member palindrome (e.g. 'eppe' in 'bell pepper')
```{r}
grep(pattern = '(.)(.)\\2\\1', dataStrings, value = TRUE, ignore.case = TRUE)
```

* ### (..)\1 ->  
(..) = Find two consecutive characters as a group. '\1' = let the group repeat again ->
Find repeating pairs of characters
```{r}
grep(pattern = '(..)\\1', dataStrings, value = TRUE, ignore.case = TRUE)
```



* ### (.).\\1.\\1 ->
'(.)' = A character, '.\'= 1 character of something else, '\1'= that 1st character again, '.\'= 1 character of something else, '\1'= that first character yet again. 

```{r}
dataStrings <- c( dataStrings, 'banana')
grep(pattern = '(.).\\1.\\1', dataStrings, value = TRUE, ignore.case = TRUE)
```

* ### (.)(.)(.).*\\3\\2\\1 ->
This takes the second expression ((.)(.)\\2\\1) one step further:
'(.)(.)(.)' = find three characters '.*' = there can be anything & any number of characters inbetween '\\3\\2\\1' = then reverse the order of the 1st three characters. (e.g 123badpassword321)

<br><br>

----------------------------------------------------------------------------
<br><br>  


## 4. Construct regular expressions to match words that:
<font size="4">

* Start and end with the same character.
* Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
* Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)


For this exercise I will use the feature 'country.name.de' from the dataframe 'codelist' from the 'countrycode' library to test my expressions
```{r, message=F}
library( countrycode )
```
```{r}
countryNames = tolower( codelist$country.name.de ) 
#i changed to lower to make the first expression a simpler case
head( countryNames )
```

* ### Start and end with the same character.
```{r}
grep(pattern = '^([A-Za-z]).*\\1$', countryNames, value = TRUE, ignore.case = TRUE)
```

* ### Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
```{r}
grep(pattern = '([A-Za-z][A-Za-z]).*\\1', countryNames, value = TRUE, ignore.case = TRUE)
```

* ### Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

```{r}
grep(pattern = '([A-Za-z]).*\\1.*\\1', countryNames, value = TRUE, ignore.case = TRUE)
```


